Predicting Covid Data

### 2022-12-05 austins_sim.rmd
added montecarlo sim

```{r}
library(tidyverse)
library(dplyr)
```

@@

```{r}
names(data)
china <- data %>% filter(location == "China", date < "2020-03-15")
```

```{r}
cases <- ggplot(china, aes(x = date, y = total_cases)) + geom_line() + labs(title = "Total Cases in China", x = "Date", y = "Total Cases")
fatalities <- ggplot(china, aes(x = date, y = total_deaths)) + geom_line() + labs(title = "Total Deaths in China", x = "Date", y = "Total Deaths")
```

```{r}
cases
fatalities
```

>>>>

```{r}
library(tidyverse)
library(dplyr)
library(pracma)
library(minpack.lm)
```

@@

AUSTRALIA SIMULATION
```{r}
australia <- data %>% filter(location == "Australia", date > "2020-06-01", date < "2020-12-19")
australia <- subset(australia, select = c(date, total_cases, total_deaths))
australia <- australia %>% mutate(days = 1:n())
```

```{r}
minimizedDays <- 70
australiaPre <- australia %>% filter(days < minimizedDays)
cases_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
print(cases_Australia)
print(fatalities_Australia)
```

```{r}
x <- 0:minimizedDays

gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
cf <- summary(nlsLM(total_cases ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=10000,b=500,c=.02,d=1)))$parameters
ff <- summary(nlsLM(total_deaths ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=100,b=100,c=.01,d=1)))$parameters

cases_function_australia <- function(x) cf[1]+cf[2]*erf(cf[3]*x-cf[4])
fatalities_function_australia <- function(x) ff[1]+ff[2]*erf(ff[3]*x-ff[4])

cases_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
print(cases_pre_plot_Australia)
print(fatalities_pre_plot_Australia)
```

```{r}
x <- 0:200

# cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
# fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)

cases_plot_Australia <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")

print(cases_plot_Australia)
print(fatalities_plot_Australia)
```

```{r}
#Montecarlo Simulation
set.seed(400)
n <- dim(australia)[1]
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

mc_data_ca <- c()
for (i in 1:n) {
    mc_data_ca <- c(mc_data_ca, australia$total_cases[i]*gau_sample[,i])
}
mc_data_ca <- matrix(mc_data_ca, nrow = m, ncol = n)
mc_data_fa <- c()
for (i in 1:n) {
    mc_data_fa <- c(mc_data_fa, australia$total_deaths[i]*gau_sample[,i])
}
mc_data_fa <- matrix(mc_data_fa, nrow = m, ncol = n)

day_of_flex_ca <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_ca[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=15000,b=9000,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_ca[i] <- min(x[temp < 0])
}
print(mean(day_of_flex_ca))
print(sd(day_of_flex_ca))
```

```{r}
day_of_flex_fa <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_fa[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=300,b=200,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_fa[i] <- min(x[temp < 0])
}
print(mean(day_of_flex_fa))
print(sd(day_of_flex_fa))
```

```{r}
print(fatalities_plot_Australia + geom_vline(xintercept = mean(day_of_flex_fa), color="#1b95ff", size=1.5, linetype="dashed"))
print(cases_plot_Australia + geom_vline(xintercept = mean(day_of_flex_ca), color="#d33636", size=1.5, linetype="dashed"))
```

### 2022-12-05 main.rmd
added montecarlo sim

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# useful libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(pracma)
# read covid data
data <- read_csv("owid-covid-data.csv")
```

## The Monte Carlo Study:

**Mathematical prediction of the time evolution of the COVID-19 pandemic in Italy by a Gauss error function and Monte Carlo simulations**

<br>

*Author: Ignazio Ciufolini, Antonio Paolozzi*

## Introduction : Ciufolini and Paolozzi's Study

* In this study by Ignazio Ciufolini and Antonio Paolozzi, they use a Gauss Error function that approximates the distribution of the cumulative positive cases of COVID-19 in China
  * They have then use the function to approximate the number of cumulative positive cases in Italy
  * Then performed a 150 Monte Carlo simulations to predict the peak and valley of the number of daily positive cases and fatalities
* This study do not take into account factors such as daily nasopharyngeal swabs, medical, social distancing, etc.

## Cont.

@@

* create a random matrix : m X n
  * m = number of random outcomes (*chosen to be 150*)
  * n = number of observed days (*n = 1,...,j*)
* each # in the matrix is a Gaussian distribution with mean = 1 and sigma = 0.1
  * multiply each column (*j*) by the # of total cases (or deaths) that corresponds to each day *j* **#FIXME#**
- add more...

@@

## Australia Simulation
```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(pracma)
data <- read_csv("owid-covid-data.csv")
australia <- data %>% filter(location == "Australia", date > "2020-06-01", date < "2020-12-19")
australia <- subset(australia, select = c(date, total_cases, total_deaths))
australia <- australia %>% mutate(days = 1:n())
minimizedDays <- 70
australiaPre <- australia %>% filter(days < minimizedDays)
cases_Australia <- ggplot(australiaPre, aes(x = date, y = total_cases)) + 
    geom_line() + 
    labs(title = "Total Cases in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Date", y = "Total Cases")
fatalities_Australia <- ggplot(australiaPre, aes(x = date, y = total_deaths)) +
    geom_line() + 
    labs(title = "Total Deaths in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Date", y = "Total Deaths")
x <- 0:minimizedDays
cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)
cases_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_line() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=2, linetype="dotted") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 1)) +
    labs(title = "Total Cases in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_line() + 
    stat_function(fun=fatalities_function_australia, color="#3333e2", size=2, linetype="dotted") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 1)) +
    labs(title = "Total Deaths in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
x <- 0:200
cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)
cases_plot_Australia <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_line() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 1)) +
    labs(title = "Total Cases in Australia from 6/1/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_line() + 
    stat_function(fun=fatalities_function_australia, color="#3333e2", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 1)) +
    labs(title = "Total Deaths in Australia from 6/1/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")
```
```{r, echo=FALSE}
# print(cases_Australia)
# print(fatalities_Australia)
print(cases_pre_plot_Australia)
# print(fatalities_pre_plot_Australia)
# cases_plot_Australia
# print(fatalities_plot_Australia)
```

## Australia results

```{r, echo=FALSE}
# print(cases_Australia)
# print(fatalities_Australia)
# print(cases_pre_plot_Australia)
# print(fatalities_pre_plot_Australia)
cases_plot_Australia
# print(fatalities_plot_Australia)
```

@@

## Works Cited
- Link to study:
- I. Ciufolini, A. Paolozzi, Prediction of the time evolution of the Covid-19 Pandemic in Italy by a Gauss
Error Function and Monte Carlo simulations. Submitted to BioRxiv on 03.26.2020 and transferred on
03.27.2020 to MedRxiv. https://doi.org/10.1101/2020.03.27.20045104
- Link to covid data:
- https://covid19.who.int/
- source to find the second derivative of fitted function:
- https://proofwiki.org/wiki/Derivative_of_Error_Function

>>>>

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# useful libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(pracma)
library(minpack.lm)
# read covid data
data <- read_csv("owid-covid-data.csv")
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

## Introduction

- In this study by Ignazio Ciufolini and Antonio Paolozzi, they use a Gauss Error function that approximates the distribution of the cumulative positive cases of COVID-19 in China
- They then use the function to approximate the number of cumulative positive cases in Italy
- Then performed 150 Monte Carlo simulations to predict the peak and valley of the number of daily positive cases and fatalities
- This study does not take into account factors such as daily nasopharyngeal swabs, medical, social distancing, etc.

## Ciufolini and Paolozzi's Study

@@

- add more...

@@

## Initial Data for Cases in Australia

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
australia <- data %>% filter(location == "Australia", date > "2020-06-01", date < "2020-12-19")
australia <- subset(australia, select = c(date, total_cases, total_deaths))
australia <- australia %>% mutate(days = 1:n())
minimizedDays <- 70
australiaPre <- australia %>% filter(days < minimizedDays)
cases_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_Australia)
```

## Initial Data for Fatalities in Australia

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_Australia)
```

## Fitting Gauss Error Function to Cases in Australia

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
x <- 0:minimizedDays

gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
cf <- summary(nlsLM(total_cases ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=10000,b=500,c=.02,d=1)))$parameters
ff <- summary(nlsLM(total_deaths ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=100,b=100,c=.01,d=1)))$parameters

cases_function_australia <- function(x) cf[1]+cf[2]*erf(cf[3]*x-cf[4])
fatalities_function_australia <- function(x) ff[1]+ff[2]*erf(ff[3]*x-ff[4])

cases_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_pre_plot_Australia)
```

## Fitting Gauss Error Function to Fatalities in Australia

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_pre_plot_Australia)
```

## Applying the function to the whole cases dataset

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
x <- 0:200

# cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
# fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)

cases_plot_Australia <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia)
```

## Applying the function to the whole fatalities dataset

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia)
```

## Using MonteCarlo Simulation to predict the day of flex

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Montecarlo Simulation
set.seed(400)
n <- dim(australia)[1]
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

mc_data_ca <- c()
for (i in 1:n) {
    mc_data_ca <- c(mc_data_ca, australia$total_cases[i]*gau_sample[,i])
}
mc_data_ca <- matrix(mc_data_ca, nrow = m, ncol = n)
mc_data_fa <- c()
for (i in 1:n) {
    mc_data_fa <- c(mc_data_fa, australia$total_deaths[i]*gau_sample[,i])
}
mc_data_fa <- matrix(mc_data_fa, nrow = m, ncol = n)

day_of_flex_ca <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_ca[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=15000,b=9000,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_ca[i] <- min(x[temp < 0])
}
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
day_of_flex_fa <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_fa[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=300,b=200,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_fa[i] <- min(x[temp < 0])
}
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
print(mean(day_of_flex_fa))
print(sd(day_of_flex_fa))
print(mean(day_of_flex_ca))
print(sd(day_of_flex_ca))
```

- The mean of the day of flex for cases is 60.22 days and the standard deviation is 0.77 days.
- The mean of the day of flex for fatalities is 81.4 days and the standard deviation is 0.69 days.

## Applying the day of flex indicator to the data (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia + geom_vline(xintercept = mean(day_of_flex_ca), color="#d33636", size=1.5, linetype="dashed"))
```

## Applying the day of flex indicator to the data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia + geom_vline(xintercept = mean(day_of_flex_fa), color="#1b95ff", size=1.5, linetype="dashed"))
```

@@

## Works Cited
- Link to study:
- I. Ciufolini, A. Paolozzi, Prediction of the time evolution of the Covid-19 Pandemic in Italy by a Gauss
Error Function and Monte Carlo simulations. Submitted to BioRxiv on 03.26.2020 and transferred on
03.27.2020 to MedRxiv. https://doi.org/10.1101/2020.03.27.20045104
- Link to covid data:
- https://covid19.who.int/
